---
title: "Exercise18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 18

Prostate-specific antigen (PSA) is an enzyme whose level is used as a (controversial) indicator of prostate cancer; however, a high level may also indicate other prostate diseases. The Prostate data set examines the correlation between PSA levels and other clinical measures in men who were about to receive a radical prostatectomy. The data set is included in the R package lasso2 and can be loaded as follows:



```{r}
data(Prostate, package = "lasso2")
summary(Prostate)

```

a) Fit a linear model to the data set using LASSO regression. Use the function cv.glmnet from the R package glmnet. Which is the optimal regularization parameter lambda ? Which coeficients are non-zero in the corresponding model?

```{r}
library("glmnet")
X <- as.matrix(Prostate[,-ncol(Prostate)])
set.seed(1234)
prostate.lasso.cv <- cv.glmnet(X, Prostate$lpsa)
plot(prostate.lasso.cv)

```
Lambda is very low at its minimum.
```{r}
prostate.lasso.cv$lambda.min
```
How to chose the lambda: https://www.youtube.com/watch?v=8oEZkHqf_Rk&list=PLOg0ngHtcqbPTlZzRHA2ocQZqB1D_qZ5V&index=36&t=0s 
Here they use cross-validation (CV)

It is better to take the biggest lambda within 1sd of the smallest MSE

```{r}
prostate.lasso.cv$lambda.1se
```

Here are the coefficients: 

```{r}
coef(prostate.lasso.cv)

```
b

b) Fit a linear model using ridge regression. Use again the function cv.glmnet, this time with
parameter alpha = 0. Which is the optimal regularization parameter lambda here?


```{r}
set.seed(1234)
prostate.ridge.cv <- cv.glmnet(X, Prostate$lpsa, alpha = 0)
plot(prostate.ridge.cv)
```


Again to find the one standard error rule lambda: 

```{r}
prostate.ridge.cv$lambda.1se
```
```{r}
coef(prostate.ridge.cv)
```



c Plot the LASSO and ridge traces of the models from a) and b) side-by-side. What are the differences, and where do they come from?

```{r}
set.seed(7)
pro.lasso <- glmnet(X, Prostate$lpsa, alpha = 1)
set.seed(7)
pro.ridge <- glmnet(X, Prostate$lpsa, alpha = 0)
par(mfrow = c(1, 2), cex = 0.5, mar = c(3, 3, 4, 1))
plot(pro.lasso, xvar = "lambda", main = "LASSO")
plot(pro.ridge, xvar = "lambda", main = "Ridge")

```

Elastic net

```{r}
set.seed(7)
pro.net.cv <- cv.glmnet(X, Prostate$lpsa, alpha = 0.5)
plot(pro.net.cv)
```

















