---
title: "Exercise11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 11


a) Read in the data set with R. Look at the structure of the data frame you get. Do you need some
manual corrections to get the data in the right format? -> beach is a factor

```{r}
rikz <- read.table("RIKZ.txt", header = T)
rikz$Beach <- as.factor(rikz$Beach)
rikz$Sample <- NULL
rikz
```


b) Fit a linear regression model using all explanatory variables. Look at the summary output of
your fit, and analyse the residuals. Which problems do you see? How could you solve them?


```{r}
rikz.fitlm <- lm(Richness ~ ., data=rikz)
summary(rikz.fitlm)
```



```{r}
library(car)
plot(fitted(rikz.fitlm), resid(rikz.fitlm), main = "Tukey Anscombe plot")
qqPlot(resid(rikz.fitlm), dist = "norm", mean = mean(resid(rikz.fitlm)), sd = sd(resid(rikz.fitlm)), main = "qqplot")
```


c) Try to improve the quality of your model from task b) by transforming some of the variables in
the RIKZ data set. Redo the model fit with the transformed variables. Does the model fit better
now?


To improve the model fit we can try applying either a log or a square root transformation, as
these lessen extreme values in end tails. However, log(0) = -inf which would seriously distort
the model. So to avoid this from happening, add 1 to all "Richness" values.


```{r}
rikz.fit2 <- lm(log(1+Richness) ~ ., data = rikz)
plot(fitted(rikz.fit2), resid(rikz.fit2), main = "Tukey Anscombe plot")
qqPlot(resid(rikz.fit2), dist = "norm", mean = mean(resid(rikz.fit2)), sd = sd(resid(rikz.fit2)), main = "qqplot")

```


We can also try with the square root transform:

```{r}
rikz.fit3 <- lm(sqrt(Richness) ~ ., data = rikz)
plot(fitted(rikz.fit3), resid(rikz.fit3), main = "Tukey Anscombe plot")
qqPlot(resid(rikz.fit3), dist = "norm", mean = mean(resid(rikz.fit3)), sd = sd(resid(rikz.fit3)), main = "qqplot")
```


d) Continue with the transformation you chose in c). Perform model selection by iteratively removing
explanatory variables. Which model do you end up with?


```{r}
library(MASS)

# fit2 -> log transformed
rikz.bw <- stepAIC(rikz.fit2, direction = "backward", trace = 0)
summary(rikz.bw)
```


```{r}
# fit3 -> squared transformed
rikz.bw2 <- stepAIC(rikz.fit3, direction = "backward", trace = 0)
summary(rikz.bw2)
```


