---
title: 'Series 1: Malthus and Verhulst model'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Thibault Schowing

## Classical models in biology (exercises) BL.6003

# Malthus model

Use the function “ode” to solve numerically Malthus’ model. Show the trajectory for the two following set of parameters and initial conditions.





```{r}
library(deSolve)
library(ggplot2)
library(tidyr)
#?ode
```


a) $r = 0.2$, and $N_0 = 5$

```{r}

# Parameters passed to func
p <- list(r = 0.2)

# Integrator
m <- c("ode45")

# Differential equation
#
# Params: 
#        - t: never used here but must remains. Offers the possibility 
#             to have time dependant differential equations.
#        - N: population size at which the time derivative will be evaluated. 
#             A list with one element per dimention. For Malthus it is simply 
#             a list of 1 element.
#        - p: parameters list 

f <- function(t,N,p){
  dN <- p$r * N
  return(list(c(dN)))
}

# Time stemps for the numerical solution
time_steps <- seq(0,10,0.01)

# Initial state values for the ODE system
N0 <- 5

# Run the "ode" function
out <- ode(y = N0, times = time_steps, func = f, parms = p, method = m)
head(out)

```


### Plot of the output


```{r}
# Simple plot
#plot(out, type = "l", xlab = "Time", ylab = "Population Size")

# ggplot looks nicer
data <- as.data.frame(out)
colnames(data) <- c("Time", "Population_size")

ggplot(data, aes(x=Time, y=Population_size)) + 
  geom_line(color="blue") +
  ggtitle("Malthus model - N0 = 5 and r = 0.02") +
  xlab("Time") +
  ylab("Population Size")
  
```






b) $r = -2$, and $N_0 = 5$

```{r}

# Parameters passed to func
p <- list(r = -2)

# Integrator
m <- c("ode45")

# Differential equation
#
# Params: 
#        - t: never used here but must remains. Offers the possibility 
#             to have time dependant differential equations.
#        - N: population size at which the time derivative will be evaluated. 
#             A list with one element per dimention. For Malthus it is simply 
#             a list of 1 element.
#        - p: parameters list 

f <- function(t,N,p){
  dN <- p$r * N
  return(list(c(dN)))
}

# Time stemps for the numerical solution
time_steps <- seq(0,10,0.01)

# Initial state values for the ODE system
N0 <- 5

# Run the "ode" function
out <- ode(y = N0, times = time_steps, func = f, parms = p, method = m)
head(out)

```




```{r}
data <- as.data.frame(out)
colnames(data) <- c("Time", "Population_size")

ggplot(data, aes(x=Time, y=Population_size)) + 
  geom_line(color="blue") +
  ggtitle("Malthus model - N0 = 5 and r = -2") +
  xlab("Time") +
  ylab("Population Size")
  
```






















# Verhulst model

a) $r = 2$, $\alpha = 0.1$, and $N_0 = 1$.

b) $r = 2$, $\alpha = 0.1$, and $N_0 = 30$.

c) $r = -2$, $\alpha = 0.1$, and $N_0 = 50$.


```{r}
p_a <- list(r = 2, alpha = 0.1)
p_b <- list(r = 2, alpha = 0.1)
p_c <- list(r = -2, alpha = 0.1)


# Verhulst differential
f_verhulst <- function(t,N,p_a){
  dN <- N * (p_a$r - (p_a$alpha * N))
  return(list(dN))
}

time_steps <- seq(0,10,0.05)

N0_a <- 1
N0_b <- 30
N0_c <- 50

out_a <- ode(y = N0_a, times = time_steps, func = f_verhulst, 
             parms = p_a, method = c("ode45"))
out_b <- ode(y = N0_b, times = time_steps, func = f_verhulst, 
             parms = p_b, method = c("ode45"))
out_c <- ode(y = N0_c, times = time_steps, func = f_verhulst, 
             parms = p_c, method = c("ode45"))


```




Prepare the Data frame for ggplot2

```{r}
data <- as.data.frame(out_a)
data["2"] <- out_b[,2]
data["3"] <- out_c[,2]

colnames(data) <- c("Time", "Pop_a_size", "Pop_b_size", "Pop_c_size")
```

Plot of the three population trajectories 
```{r}
test_data <- data
test_data %>%
  gather(key,value, Pop_a_size, Pop_b_size, Pop_c_size) %>%
  
  ggplot(aes(x=Time, y=value, colour=key)) +
  geom_line(alpha=0.55, size=0.8) +
  geom_point(mapping=(aes(x=0,30)), color="green", size=2) +
  geom_point(mapping=(aes(x=0,50)), color="blue", size=2) +
  geom_point(mapping=(aes(x=0,1)), color="red", size=2) +
  ggtitle("Verhulst model ") +
  xlab("Time") +
  ylab("Population Size") +
  xlim(c(0,10))
```

- Populations:

a) $r = 2$, $\alpha = 0.1$, and $N_0 = 1$.

b) $r = 2$, $\alpha = 0.1$, and $N_0 = 30$.

c) $r = -2$, $\alpha = 0.1$, and $N_0 = 50$.



