---
title: "Exercise10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 10

[Graded] Researchers measured the yield of 5 crop species treated with 4 fertilizers. You find the measurements in the file fertilizer.dat on ILIAS.


a) Read in the data set and draw an interaction plot.
R hint: since the factor levels are not encoded as strings, you first have to explicitely convert the variables CropSpecies and Fertilizer to factors.



```{r}
library(car)
data <- read.table("fertilizer.dat", header=TRUE)
data$Fertilizer <- as.factor(data$Fertilizer)
data$CropSpecies <- as.factor(data$CropSpecies)

interaction.plot(data$Fertilizer, data$CropSpecies, data$Yield)

```


We see that the five species act the same with the different fertilizer (they all increase similarly). There might be an interaction. 


b) Fit an ANOVA model without interaction to the data set. Analyse the residuals. Does the model fit well? If not, can you think of a transformation that could make the fit better?


```{r}
data.fit <- lm(Yield ~ Fertilizer + CropSpecies, data = data)
plot(fitted(data.fit), resid(data.fit), main = "Tukey Anscombe plot")
qqPlot(resid(data.fit), dist = "norm", mean = mean(resid(data.fit)), sd = sd(resid(data.fit)), main = "qqplot")


```




The two plots demonstrate a poor fit. In the Tukey Anscombe plot we can clearly see a pattern between the residuals and the fitted values that should be randomly distributed if the model assumptions were right. The qqPlot is right skewed which suggests that we should try a log transform.

```{r}
data.fit2 <- lm(log(Yield) ~ Fertilizer + CropSpecies, data = data)
plot(fitted(data.fit2), resid(data.fit2), main = "Tukey Anscombe plot")
qqPlot(resid(data.fit2), dist = "norm", mean = mean(resid(data.fit2)), sd = sd(resid(data.fit2)), main = "qqplot")
```


It looks much better!

c) Add an interaction term to the model from task b). Is there a significant interaction?



```{r}

data.fit3 <- lm(log(Yield) ~ Fertilizer*CropSpecies, data = data)

#Note: if we try to compare with the first model, R will show an error because we need to have the log() in both.
anova(data.fit2, data.fit3)
```

Value of the F statistic is not within the range of rejection, hence the interaction is not significant.

d) Beginning from the model from task c), perform a model selection by remove non-significant terms. In which order do you have to test the terms? With which model do you end up?


As we saw in point c, the model with the interaction term is not significant. So, we can already remove that and start again with the following model: 


```{r}
data.fit2 <- lm(log(Yield) ~ Fertilizer + CropSpecies, data = data)
```


Next we need to check the significance of Fertilizer and CropSpecies:


```{r}
data.fitFertilizer <- lm(log(Yield) ~ Fertilizer, data = data)
summary(data.fitFertilizer)
data.fitCropSpecies <- lm(log(Yield) ~ CropSpecies, data = data)
summary(data.fitCropSpecies)
```

From this we conclude that both variables are significant. Therefore, our best model is: 
log(Yield) ~ Fertilizer + CropSpecies
