---
title: "Exercise3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Compute an ordinary linear regression of I versus C. Does the income depend on the number of cows?
```{r}
farm <- read.table("farm.dat", header = TRUE)
farm.fit <- lm(Dollar ~ cows, data=farm)
summary(farm.fit)

```


The p-value of the regression coefficient is very small (***) thus shows that the Income explains very well the number of cows. 


Give the confidence intervals for the expected income without any cows, with 20 cows, and with C = 8.85 cows.

```{r}
predict(farm.fit, newdata=data.frame(cows=c(0,20,8.85)), interval="confidence")
```

Compute an ordinary linear regression of I versus A and a multiple linear regression of I versus A and C. Also compute the correlation between A and C. Finally, based on your results, explain the differences between the three regression models.


First, we fit the Income vs Acres
```{r}
fitAI <- lm(Dollar~acres, data=farm)
summary(fitAI)
```


There is no significance (*) between the income and the acres. We add Cows as a covariate of Acres (Does income explain Acres + Cows?).

```{r}
fitAIC <- lm(Dollar~acres+cows, data=farm)
summary(fitAIC)
```


As covariate, both Acres and Cows are significant.



```{r}
corCA <- cor(farm$cows, farm$acres); corCA

```

We can see that acres and cows have a negative corelation and that cows and income have a positive one. Acres does not explain the income (big farm with no cows for examples) but when put together

To have a nice overview:

```{r, echo=FALSE}
library(ggplot2)
library(gridExtra)


p1 = ggplot(data = farm,aes(x=Dollar,y=cows, color=acres)) + 
  geom_point() +
  geom_smooth(method = lm) +
  theme_classic()

p2 = ggplot(data = farm,aes(x=cows,y=acres, color=Dollar)) + 
  geom_point() +
  geom_smooth(method = lm) +
  theme_classic()

p3 = ggplot(data = farm,aes(x=acres,y=Dollar, color=cows)) + 
  geom_point() +
  geom_smooth(method = lm) +
  theme_classic()



grid.arrange(p1,p2,p3,nrow = 3)

```

