---
title: "Exercise15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 15

```{r}

library(lme4)
data(Pastes)
```


Assume you would have forgotten your knowledge on random effects and want to analyse the
data set with an one- or two-way ANOVA. How do you have to specify the model, and what are
the problems of such an approach?

```{r}
dim(Pastes)
```

```{r}
head(Pastes)
```
```{r}
summary(Pastes)
```



```{r}
fit1 <- lm(strength ~ sample, data=Pastes)
fit2 <- lm(strength ~ batch, data=Pastes)
fit3 <- lm(strength ~ batch+sample, data=Pastes)
fit4 <- lm(strength ~ batch+cask, data=Pastes)
```

Problems: 

* too many variables with small group sizes.
* missing variables for ex in fit3
* coeff for single factor only 
* casks (fût) are assumed to be the same between the different batch even if it is probably just a label (fixed effect of "a" which means nothing)

We it shows the that fitting Anova model with fixed effects to data where random effects are present and have to be taken in account. 











Assuming you have not forgotten anything on random effects. Fit a two-way random effects
model to the data set. Which model formula is appropriate? Which remarkable feature does the
data have?

We can model this in different manner to have the "two-way": 

batch & cask, batch & sample and cask & sample. But even with random effects, cask stays not practical to use (share the same label etc..) so we take batch & sample. 

It is also worth noting that this: experiment has a nested design, as first the batch is randomly selected and then from each selected batch the casks are randomly taken. 
This limits the choice of indices available in the formula for this two-way ANOVA with random effects: 

we have i = 1...g batches, j = 1, ..., n casks and k = 1,2 assays 


```{r}
pastes.fit <- lmer(strength ~ 1 + (1|batch) + (1|sample), data = Pastes)
summary(pastes.fit)
```


Does the strength significantly deviate between different casks? And between different batches?



```{r}
set.seed(42)
confint(pastes.fit)
```

Here we can see that zero is included in the interval


```{r}
pastes.fit2 <- lmer(strength ~ 1 + (1|cask) + (1|sample), data = Pastes)
set.seed(42)
confint(pastes.fit2)
```






