---
title: "Part2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Characterize the population structure with individual-based methods


```{r}
# read the genotype matrix
geno <- as.matrix(read.table("./Henn_et_al_data/Hennetal_genotypeMatrix.geno", header=T, stringsAsFactors = F, na.strings = "NA"))
# check if we read genotype matrix correctly
# str(geno) # matrix with nsites rows and nind columns

# ID of individuals correspond to column names
inds <- colnames(geno) 

# Get the index of the individuals from each population
pop.names <- c("San","Mbuti","Mozabite","Pathan","Cambodian","Yakut","Maya")
sample_inds <- list() # list to save the index of individuals belonging to each pop
for(i in 1:length(pop.names)) {
  # use the function which() to get the index
  # of individuals whose initial characters of the name are equal to pop.names
  sample_inds[[i]] <- which(substr(inds, 1, nchar(pop.names[i]))==pop.names[i])
}
# check the result
# str(sample_inds)

# Get a vector with the population for each individual
popinds <- inds # just initiliaze to the individual labels
for(i in 1:length(pop.names)) {
  popinds[sample_inds[[i]]] <- pop.names[i]
}
# check the result
# str(popinds)
```


```{r}
# load LEA package
library("LEA")
 
# create tag for filename with results
filename_tag <- "HumanData"

# Need to create the input file (geno SNMF file)
genoSNMF <- geno
# replace missing data by "9"
genoSNMF[is.na(genoSNMF)]=9
# save the file with the genotypes using 9 as missing data
write(t(genoSNMF),file=paste(filename_tag, "_sNMF.geno",sep=""), ncolumns = ncol(genoSNMF), sep="")

# Perform PCA with LEA package
# Create a pcaProject object: pc.
pc <- pca(paste(filename_tag, "_sNMF.geno",sep=""), scale = FALSE, center=TRUE)
```



```{r}
#######################
# Display Information #
#######################

# Display information about the analysis.
#show(pc)

# Summarize the analysis.
#summary(pc)

#####################
# Graphical outputs #
#####################
library(RColorBrewer)
mycols <- brewer.pal("Paired", n=length(pop.names))

# get the explained percentage
perc_exp <- pc$eigenvalues/sum(pc$eigenvalues, na.rm=T)
barplot(perc_exp, ylab="Percentage variance explained", xlab="PCs", beside=T)
```



```{r}
# Look at the resulting eigenvectors
indcolor <- as.numeric(factor(popinds)) # set the color for each individual
# Plot the first PCs
par(mfrow=c(2,3))
for(i in 1:3) {
  for(j in (i+1):4) {
    plot(pc$projections[,i],pc$projections[,j], col=mycols[indcolor], pch=16, xlab=paste("PC", i, "%var=", perc_exp[i]), ylab=paste("PC", j, "%var=", perc_exp[j]),)
    if(i==1 & j==2) {
    legend("top", pop.names, col=mycols[factor(pop.names)], pch=16)      
    }
    
  }
}
```


```{r}
#############################
# Perform Tracy-Widom tests #
#############################

# Perfom Tracy-Widom tests on all eigenvalues.
# Create file: genotypes.tracyWidom - tracy-widom test information, 
#          in the directory genotypes.pca/.
#tw = tracy.widom(pc)

# BUG I FOUND: Might need to check the output file from PCA, to remove scientific notation values... Open the file and change it in a text editor.

# Display the p-values for the Tracy-Widom tests. 
#par(mfrow=c(1,1))
#plot(log10(tw$pvalues))

# How many PCs are significant?
```

We can si (when it works) that the four first components are significant. We can also see this on the graph.


## 3.2. Infer individual ancestry coefficients with sparse Nonlinear Matrix Factorization (sNMF)

second method


## Exercise 2

- Use sNMF to analyse the Henn et al (2015) data

- What is the best K to explain the data?

- How many populations do we have in this dataset?

- Can you detect any migrant individual?




```{r}
# load LEA package
library("LEA")
 
# Need to create the input file (geno SNMF file)
genoSNMF <- geno
# replace missing data by "9" -> old but strange tradtition
genoSNMF[is.na(genoSNMF)]=9
# save the file
write(t(genoSNMF),file=paste(filename_tag, "_sNMF.geno",sep=""), ncolumns = ncol(genoSNMF), sep="")
# run sNMF from K values 1 up to 5
obj.at = snmf(paste(filename_tag, "_sNMF.geno",sep=""), K=1:5, ploidy=2, entropy=T, CPU=3, project="new", repetitions=1)

# look at what is in obj.at
str(obj.at)

# Find the index of the end of each population
npops <- length(pop.names)
indexpop <- numeric(npops)
for(i in 1:npops){
  indexpop[i] <- max(sample_inds[[i]])
}

pdf("AdmixtureProportions.pdf", width=10, height=6)
# Plot the results for each K value and the cross-entropy
par(mfrow=c(1,1))
for(k in 2:5) {
  mycols=brewer.pal(k, "Paired")
  # the Q() function gets the ancestry coefficients for each K value
  bp=barplot(t(Q(obj.at, K=k)), horiz = F, beside = F, col = mycols, space=0.01, border=NA)
  # add lines dividing the populations
  abline(v=indexpop, col=1, lwd=2)
  # add the population names below each area
  axis(side=1, at=indexpop-2, labels=pop.names)
}
# Plot the cross-entropy
par(mfrow=c(1,1))
plot(obj.at)
dev.off()

# What is the best K-value? 
```

For each value of k the prog. calculate the ancestry ...
Check the created PDF.

Each bar is one individual
Looking at the four different K (acncestral pop) we can see that Mbuti and San are assigned to 1 ancestral population and the rest to 2. We can see that Mozabite and Pathan have the two ancestries at different proportions. 

In the second graph with K=3, we see that Mbuty and san are assigned to different ancestry and the rest of pop are assighed to a third ancestry. 

The best K (last graph, cross enthropy): the lowest is for k=2 -> meaning that it's the best explanation. 2 ancestral pop: mbuti/san to the rest and Mozabite and pathan that have mixed ancestry. 

TO DISCUSS


- Do the PCA and Ancestry proportions analysis indicate the same population structure?


Yes. On PCA, first axix separates san/mbuti with Mozambite intermediate. On the second: pathan and third: san/mbuti. 




- Based on these results, what is the human population structure?



- What are the likely demographic events that lead to this structure?


As we said, it's linked to the OOA. 


- Is there any cryptic population structure within each sampled location?

Each pop is homogenous, even the mixed ones have the same pattern of ancestry

- Is there evidence of hybrid (admixed) individuals between different populations?

Yes, Mozabite and pathan

- Is there evidence of shared ancestry? That is, two populations that descend from the same ancestral population?


Yes, All the pops on the right (light blue ancestry) or left (dark blue ancestry). 


































































